---
title: "Jackal Interactive Map"
format: html
editor: visual
---

## Exploring the distribution of two jackal species in Southern Africa

This is an exercise to practice using GIS. I will be using point observation data from iNaturalist to plot the distribution of the two jackal species found in Southern Africa: the black-backed jackal (Lupulella mesomalas) and the side-striped jackal (Lupulella adusta).

## Preparing the dataset

Call in the necessary packages

```{r}
packages <- c("rinat", "sf", "tidyverse", "ggspatial", "htmltools", "mapview", "leafpop")
lapply(packages, library, character.only = TRUE)
```

Call necessary data from iNaturalist for both jackal species.

```{r}
lm <- get_inat_obs(taxon_name = "Lupulella mesomelas",
                   bounds = c(-36, 10, -11, 42),
                   maxresults = 1000)
la <- get_inat_obs(taxon_name = "Lupulella adusta",
                   bounds = c(-36, 10, -11, 42),
                   maxresults = 1000)
```

Combine into one dataframe

```{r}
lup <- rbind(lm,la)
```

Filter observations by attribute criteria

```{r}
lup <- lup %>% filter(positional_accuracy<46 & 
                        latitude<0 &
                        !is.na(latitude) &
                        quality_grade == "research")
```

Create new column so observations recorded as a subspecies just have genus and species name

```{r}
lup <- lup %>%
  mutate(species = word(scientific_name, 1, 2))
```

Make the dataframe a spatial object of class = "sf" and set coordinate reference system (CRS)

```{r}
lup <- st_as_sf(lup, coords = c("longitude", "latitude"), crs = 4326)
```

Plot interactive map with html links to iNaturalist records

```{r}
llup <- lup %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))

mapview(lup, 
        zcol = "species",
        col.regions = c("darkcyan", "orangered3"),
        legend = TRUE,
        layer.name = "Jackal species",
        popup = 
          popupTable(llup,
                     zcol = c("user_login", "click_url")))
```
